---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r, echo = TRUE}
data <- read.csv("activity.csv")
# summary(data)
data$date <- as.Date(data$date, format="%Y-%m-%d")
```

## What are the total number of steps per day?
```{r, echo = TRUE}
steps_per_day <- aggregate(steps ~ date, data, sum)
hist(steps_per_day$steps, xlab = 'Number of Steps', main = 'Histogram of Total Number of Steps per Day')
```

## What is mean total number of steps taken per day?

```{r, echo = TRUE}
mean_steps <- mean(steps_per_day$steps)
median_steps <- median(steps_per_day$steps)

print(paste("The mean steps per day is ",mean_steps, " and the median is ", median_steps,".", sep = ""))

```

## What is the average daily activity pattern?
```{r, echo = TRUE}
steps_per_interval <- aggregate(steps ~ interval, data = data, FUN = mean, na.rm = TRUE)
plot(steps_per_interval$interval,steps_per_interval$steps,xlab = 'Interval',type = "l")
```


## What is the intervale with the greatest average of steps?
```{r, echo = TRUE}
interval_with_max_steps <- steps_per_interval[which.max(steps_per_interval[,2]),1]
print(paste("The interval with the most steps is interval ",interval_with_max_steps,'.',sep = ''))
```
## Imputing missing values
### Number of missing values
```{r, echo=TRUE}
number_of_missing <- sapply(data, function(x) sum(is.na(x)))
print(number_of_missing)
```

### Replace Nan values with the median based on the interval
```{r, echo=TRUE}

library(dplyr)
library(tidyr)
clean_data <- data %>% 
  group_by(date) %>% 
  mutate_at(vars(steps), ~replace_na(., median(., na.rm = TRUE)))

summary(clean_data)
```
### Histogram of total steps per day
```{r, echo = TRUE}
steps_per_day_clean <- aggregate(steps ~ date, clean_data, sum)
hist(steps_per_day_clean$steps, xlab = 'Number of Steps', main = 'Histogram of Total Number of Steps per Day')
```

### The mean and median of cleaned data
```{r, echo = TRUE}
mean_steps_clean <- mean(steps_per_day_clean$steps)
median_steps_clean <- median(steps_per_day_clean$steps)

print(paste("The mean steps per day is ",mean_steps_clean, " and the median is ", median_steps_clean,".", sep = ""))

```

## Are there differences in activity patterns between weekdays and weekends?
### Determine if day is on the weekend.
```{r, echo = TRUE}
clean_data$day_of_week <- weekdays(clean_data$date, abbreviate = TRUE)
clean_data$day_of_week <- ifelse(clean_data$day_of_week=="Sun"|clean_data$day_of_week=="Sat","Weekend","Weekday")
```
### Plot data

```{r, echo = TRUE}
steps_for_day_of_week <- aggregate(steps ~ interval+day_of_week, data = clean_data, FUN = mean, na.rm = TRUE)
day_of_week <- split(clean_data,f=clean_data$day_of_week) 

library(ggplot2)

ggplot(steps_for_day_of_week, aes(x=interval, y=steps)) +
    geom_line() +
    facet_wrap(~ day_of_week, nrow=2, ncol=1)
```

